---
title: ''
output: 
    bookdown::word_document2: 
      reference_docx: word-styles-reference.docx #make sure this line is indented! see this guide for formatting: https://rmarkdown.rstudio.com/articles_docx.html
bibliography: references.bib
csl: new-phytologist.csl

nocite: |
  @harris_updated_2014; @jung_exploiting_2006; @besnard_mapping_2021; @besnard_global_2024
---


**Title:** Informing forest carbon inventories under the Paris Agreement using ground-based forest monitoring data


**Authors:**

Kristina J. Anderson-Teixeira^1,2^*

Valentine Herrmann^1^

Madison Williams^1^

Teagan Tinuviel^3^ (ORCID: 0000-0002-8828-9640)

Rebecca Banbury Morgan^1,4,5^

Ben Bond-Lamberty^6^

Susan Cook-Patton^7^



**Affiliations:**

1. Center for Conservation Ecology, Smithsonian's National Zoo & Conservation Biology Institute, Front Royal, VA, United States

2. Forest Global Earth Observatory, Smithsonian Tropical Research Institute, Panama, Republic of Panama

3. Smithsonian-Mason School of Conservation, Front Royal, VA, United States

4. School of Geography, University of Leeds, Leeds, United Kingdom

5. School of Biological Sciences, University of Bristol, Bristol, United Kingdom 

6. Joint Global Change Research Institute, Pacific Northwest National Laboratory, College Park, MD, United States

7. The Nature Conservancy; Arlington VA 22203, United States

*corresponding author: teixeirak@si.edu

<!--
special issue: Anthromes and terrestrial carbon: from the deep past to net-zero
special issue description: https://www.newphytologist.org/news/view/332#submission-procedure
journal: Plants, People, Planet
author guidelines: https://nph-onlinelibrary-wiley-com.smithsonian.idm.oclc.org/hub/journal/25722611/about/author-guidelines
Manuscript Category: Methods and Techniques
Length: usually in the region of 3000–4000 words
Display Items: 6
-->

```{r eval = TRUE, echo=FALSE, warning=FALSE, results='hide'}
#Tables
table_pools =1
table_recommendations = 2

#Figures
fig_variable_mapping=1
fig_map=2
fig_climate=3
fig_histograms=4


#Appendices
appendix_ForC_updates = "Methods S1"
appendix_IPCC_definitions = "Notes S1"
table_ForCchanges = "Table S1"
table_ForCfieldmapping = "Table S2"
table_variables ="Table S3"

#READING IN FILES TO GET FIGURES/ TABLES/ STATS 
path_to_dir <- dirname(dirname(getwd())) # get path to repo 
path_to_ForC= "/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db" #Krista's laptop
# path_to_ForC= "/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db" #Madison, to knit, you'll need to enter the path to ForC-db on your computer, then un-comment this line.

# MOVE FIGURE FILES 
#world map from ForC GitHub
do.call(file.remove, list(list.files(paste0(path_to_dir, "/doc/manuscript/figures_tables"), "World_Map_of_sites_with_FAO_and_IPCC_data_sent.png", full.names = TRUE))) # first remove old file
flist <- list.files(paste0(path_to_ForC, "/ForC/figures/"), "World_Map_of_sites_with_FAO_and_IPCC_data_sent.png", full.names = TRUE)
file.copy(flist, paste0(path_to_dir, "/doc/manuscript/figures_tables"))

#climate diagram from ForC GitHub
do.call(file.remove, list(list.files(paste0(path_to_dir, "/doc/manuscript/figures_tables"), "Climate_of_forC_sites_IPCC.png", full.names = TRUE))) # first remove old file
flist <- list.files(paste0(path_to_ForC, "/ForC/figures/"), "Climate_of_forC_sites_IPCC.png", full.names = TRUE)
file.copy(flist, paste0(path_to_dir, "/doc/manuscript/figures_tables"))

#Statistics
ForC_measurements <- read.csv(paste0(path_to_ForC,"/ForC/data/ForC_measurements.csv"), stringsAsFactors = FALSE, check.names = FALSE)
n_records_ForC= length(ForC_measurements[,"measurement.ID"])  # total n records in ForC

ForC_plots <- read.csv(paste0(path_to_ForC,"/ForC/data/ForC_plots.csv"), stringsAsFactors = FALSE, check.names = FALSE)
n_plots_ForC = length(ForC_plots[,"plot.ID"]) # total n plots in ForC

ForC_sites <- read.csv(paste0(path_to_ForC,"/ForC/data/ForC_sites.csv"), stringsAsFactors = FALSE, check.names = FALSE)
n_areas_ForC = max(ForC_sites$geographic.area, na.rm = TRUE) # total n geographic areas in ForC

ForC_simplified <- read.csv(paste0(path_to_ForC,"/ForC/ForC_simplified/ForC_simplified.csv"), stringsAsFactors = FALSE, check.names = FALSE)
n_independent_records_ForC= length(ForC_simplified[,"measurement.ID"]) #suspected duplicates removed. This is the number of records fin ForC simplified.

C_variables <- read.csv(paste0(path_to_dir,"/doc/manuscript/figures_tables/C_variables.csv"), stringsAsFactors = FALSE, check.names = FALSE)
n_relevant_variables = nrow(C_variables) - 8  # n variables relevant = rows in C_variables - headers and total 
n_relevant_variables_with_no_data =  sum(C_variables$`n in ForC` == 0,  na.rm = TRUE)
n_relevant_variables_with_data= n_relevant_variables - n_relevant_variables_with_no_data


n_ind_records_EFDB_variables= C_variables[length(C_variables[,"variable"]), "n independent records in ForC"] #n independent records relevant to EFDB
n_ind_records_agb <- C_variables$`n independent records in ForC`[C_variables$`variable` == "biomass_ag"]
n_ind_records_agb_components <- C_variables$`n independent records in ForC`[C_variables$`variable` == "biomass_ag_woody"] + C_variables$`n independent records in ForC`[C_variables$`variable` == "biomass_ag_foliage"]
n_ind_records_biomass <- C_variables$`n independent records in ForC`[C_variables$`variable` == "biomass"]
n_ind_records_rootbiomass <- C_variables$`n independent records in ForC`[C_variables$`variable` == "biomass_root"] +C_variables$`n independent records in ForC`[C_variables$`variable` == "biomass_root_fine"]+C_variables$`n independent records in ForC`[C_variables$`variable` == "biomass_root_coarse"]
n_ind_records_deadwood <- C_variables$`n independent records in ForC`[C_variables$`variable` == "deadwood"] +C_variables$`n independent records in ForC`[C_variables$`variable` == "deadwood_down"]+C_variables$`n independent records in ForC`[C_variables$`variable` == "deadwood_standing"]
n_ind_records_OL <- C_variables$`n independent records in ForC`[C_variables$`variable` == "O.horizon"]
n_ind_records_TotEcosystemC <- C_variables$`n independent records in ForC`[C_variables$`variable` == "total.ecosystem_2"] 
n_ind_records_SOC <- C_variables$`n independent records in ForC`[C_variables$`variable` == "SOM / SOC"] 
n_ind_records_deltaagb <- C_variables$`n independent records in ForC`[C_variables$`variable` == "delta.agb"] 
n_ind_records_deltaOL <- C_variables$`n independent records in ForC`[C_variables$`variable` == "delta.O.horizon"] 
flux_variables_with_records <- c("NPP_woody", "ANPP_woody", "ANPP_woody_stem", "ANPP_woody_branch", "woody.mortality_ag", "BNPP_root", "BNPP_root_fine","BNPP_root.turnover_fine","BNPP_root_coarse" , "ANPP_litterfall", "NPP_litter", "R_het_litter")
n_ind_records_flux <- sum(C_variables$`n independent records in ForC`[C_variables$`variable` %in% flux_variables_with_records])

n_records_rejected_variables = 23+5  # 23 litter, 5 NEE

n_records_reviewed_or_added= C_variables[length(C_variables[,"variable"]), "n reviewed"] + n_records_rejected_variables
n_records_added = max(ForC_measurements$measurement.ID) -48956 # Madison's entries start with measurement.ID = 48956
n_records_checked = n_records_reviewed_or_added - n_records_added  # this is hard to get at, but we can at least estimate a minimum (= previously existing records in records sent to EFDB)

n_records_sent_to_EFDB =  C_variables[length(C_variables[,"variable"]), "n submitted to EFDB"] + n_records_rejected_variables  # n records sent to EFDB
n_relevant_variables_no_records_sent =  sum(C_variables$`n submitted to EFDB` == 0,  na.rm = TRUE)
n_relevant_variables_records_sent= n_relevant_variables - n_relevant_variables_no_records_sent
n_records_sent_biomass<- C_variables$`n submitted to EFDB`[C_variables$`variable` == "biomass"]
n_records_sent_agb <- C_variables$`n submitted to EFDB`[C_variables$`variable` == "biomass_ag"]
n_records_sent_agb_components <- C_variables$`n submitted to EFDB`[C_variables$`variable` == "biomass_ag_woody"] + C_variables$`n submitted to EFDB`[C_variables$`variable` == "biomass_ag_foliage"]
n_records_sent_rootbiomass <- C_variables$`n submitted to EFDB`[C_variables$`variable` == "biomass_root"] 
n_records_sent_rootbiomass_components <- C_variables$`n submitted to EFDB`[C_variables$`variable` == "biomass_root_fine"]+C_variables$`n submitted to EFDB`[C_variables$`variable` == "biomass_root_coarse"]
n_records_sent_deadwood<- C_variables$`n submitted to EFDB`[C_variables$`variable` == "deadwood"] 
n_records_sent_deadwood_components<- C_variables$`n submitted to EFDB`[C_variables$`variable` == "deadwood_down"]+C_variables$`n submitted to EFDB`[C_variables$`variable` == "deadwood_standing"]
n_records_sent_litter<- C_variables$`n submitted to EFDB`[C_variables$`variable` == "O.horizon"]
n_records_sent_soil<- C_variables$`n submitted to EFDB`[C_variables$`variable` == "SOM / SOC"] 
n_records_sent_stocks = n_records_sent_biomass+n_records_sent_agb+n_records_sent_agb_components+n_records_sent_rootbiomass+n_records_sent_rootbiomass_components+n_records_sent_deadwood+n_records_sent_deadwood_components+ n_records_sent_litter+ n_records_sent_soil
n_records_sent_delta.agb<- C_variables$`n submitted to EFDB`[C_variables$`variable` == "delta.agb"]
n_records_sent_delta.litter<- C_variables$`n submitted to EFDB`[C_variables$`variable` == "delta.O.horizon"]
n_records_sent_increments <- n_records_sent_delta.agb+n_records_sent_delta.litter
n_records_sent_fluxes <- sum(C_variables$`n submitted to EFDB`[C_variables$`variable` %in% flux_variables_with_records])
n_records_sent_fluxes_check = n_records_sent_to_EFDB-n_records_sent_stocks-n_records_sent_increments


n_records_posted_EFDB = 1068 # n records accepted and posted in EFDB (easy to determine by entering "ForC" in the "Other Properties" search field of EFDB: https://www.ipcc-nggip.iges.or.jp/EFDB/find_ef.php. Then get number from upper left box where it says "displayed records". Updated Jan 2024)
percent_forest_records = round(n_records_posted_EFDB/5644*100,0); #as of Jan 2024

n_new_variables = 11 + 2 + 2 #delta. + woody.mortality + O horizon
# get n fields modified
sampling_details <- read.csv(paste0(path_to_ForC,"/ForC/database_management_records/record_of_changes.csv"), stringsAsFactors = FALSE, check.names = FALSE)
n_fields_modified =  length(sampling_details[,"Field"]) + 4 


```

\newpage 


# Summary {-}
<!--The Summary for research articles must not exceed 250 words and should be organised using four bullet points to indicate (1) the research conducted or aims, including the rationale, (2) methods, (3) key results, and (4) the main conclusion, including the key points of discussion. It should not contain citations of other papers. For reviews, keep to the word limit, but bullet points do not apply. -->


<!-- (1) the research conducted or aims, including the rationale-->
- Humans have been influencing Earth's climate via transformative impacts on forests for millenia, and forests are now recognized as critical to climate change mitigation under the Paris Agreement.
The efficacy of climate change mitigation planning and reporting depends on quality data on forest carbon (C) stocks and changes.
The Emission Factor Database (EFDB) of the International Panel on Climate Change (IPCC) is intended to be a definitive source for such data, but needs comprehensive and well-documented data to be so.
<!-- (2) methods-->
- To facilitate submission of forest C estimates from scientific studies to EFDB, we develop and document a process for semi-automated data submission from the Global Forest C database (ForC v4.0), which is the largest compilation of ground-based forest C estimates. 
We then assess the data currently available through ForC and provide recommendations for improving forest data collection, analysis, and reporting.
<!-- (3) key results-->
- As of `r format(Sys.time(), '%B %Y')`, ForC contained ~`r n_ind_records_EFDB_variables` independent records relevant to EFDB, `r n_records_posted_EFDB` of which had been submitted and posted to EFDB. These represented `r round(n_records_posted_EFDB/5644*100,0)`% of the total EFDB records for forest land.
Records were unevenly distributed across variables<!-- (skewed towards aboveground biomass stocks)--> and geographic regions<!-- (skewed towards temperate forests)-->.
`r round((n_records_reviewed_or_added-n_records_sent_to_EFDB)/n_records_reviewed_or_added,2)*100`% of ForC records reviewed could not be submitted because the original publication lacked required information.
<!-- (4) the main conclusion, including the key points of discussion-->
- In the future, ground-based forest C estimates should target gaps in the record, and studies should ensure that they report all information necessary for inclusion in EFDB. 
Given that climate change is rapidly impacting the world’s forests, timely reporting of recent estimates will be critical to accurate forest C inventories.  

<!--suggestions from Sandro: “In the next future, forest C IPCC default values can be improved based on data populated in the EFDB, being thus of high relevance any process for the population of the EFDB with large datasets, especially to fill critical gaps in it. Given that climate change is rapidly impacting the world’s forests, timely uploading in the EFDB of recent data is critical to support quality of estimates C stock gains and losses reported in national GHG inventories.”-->


<!--
# 200 word summary submitted to Anthromes special issue (invited for full submission)
The efficacy of climate change mitigation planning and reporting depends on high quality data on forest carbon (C) stocks and changes. The Emission Factor Database (EFDB) of the International Panel on Climate Change (IPCC) is intended as a definitive source for such data. 
The Global Forest C database (ForC) is a compilation of tens of thousands of ground-based forest C estimates that is fully traceable and publicly available. Here, we develop and document a process for semi-automated submission of data from ForC into EFDB,  assess the data currently available through ForC, and provide recommendations for improving forest data collection, analysis, and reporting. 
ForC contained ~19,316 independent records relevant to EFDB, 1,438 of which had undergone necessary review and been submitted. Records were unevenly distributed across variables and geographic regions. 59% of ForC records reviewed could not be submitted because the original publication lacked information required by the IPCC. 
In the future, ground-based forest C estimates should be target gaps in the record, and studies should ensure that they report all information necessary for inclusion in EFDB. Given that climate change is rapidly impacting the world’s forests, timely reporting of recent estimates will be especially critical to accurate forest C inventories.
-->



**Keywords:**
<!--Five to eight keywords (in alphabetical order) must be given at the end of the Summary.-->

climate change, database, forest carbon, greenhouse gas inventory, International Panel on Climate Change (IPCC), natural climate solutions, nature based climate solutions


# Societal Impact Statement {-}

<!--Please provide a Societal Impact Statement to accompany your manuscript. Societal Impact Statements should outline the wider impact or relevance of the work for people, society or the planet, including any implications for policy and/or practice. This engaging 100 word statement should be written in plain language to be accessible for non-experts.-->

Human interactions with forests have shaped Earth's climate for millennia and will continue to do so as we target net-zero emission goals.
Accurately characterizing these climate impacts requires making accurate forest carbon data available for forest monitoring and planning. 
Here we develop a semi-automated process for submitting forest carbon measurements from the largest relevant scientific database to the International Panel on Climate Change's Emission Factor Database, which currently has sparse forest carbon data.
Building this bridge from scientific research to international policy is an important step towards managing forests in a net-zero motivated future.



\newpage

# Introduction {-}

Humans have been influencing Earth's climate via ecologically transformative impacts on ecosystems for >12,000 years [@ellis_people_2021;@sanderman_soil_2017; @bonan_forests_2016].
In recent decades, as both anthropogenic land transformation and climate change have accelerated, this relationship has come into increasing focus [@ipcc_climate_2019].
Deforestation and forest degradation are substantial sources of the greenhouse gas carbon dioxide (CO~2~), currently accounting for >10% of anthropogenic emissions [@friedlingstein_global_2022].
At the same time, CO~2~ uptake by remaining and regrowing forests, woodlands, and savannas has exceeded releases from deforestation and other severe disturbances, resulting in a net carbon CO~2~ sink of ~0.88 - `r round(5.8*12/44,1)` Gt C yr^-1^ [@xu_changes_2021;@harris_global_2021],
<!--@xu_changes_2021 includes all biomes with trees, estimates ~0.88 Gt C yr^-1^
-1.6 is net forest estimate in @harris_global_2021 Table 2, which includes only forests;
BBL: might cite Ruehr et al. 2023 http://dx.doi.org/10.1038/s43017-023-00456-3 here-->
that offsets an estimated `r round(.88*44/12/32,2)*100`<!--net forest estimate from xu_changes_2021 / fossil fuel & cement, (Harris et al. Table 2)--> - `r round(5.8/32,2)*100`%<!--net forest estimate / fossil fuel & cement, (Harris et al. Table 2)--> of anthropogenic CO~2~ emissions from fossil fuels and cement [@xu_changes_2021;@harris_global_2021].
The future of this important CO~2~ sink depends primarily upon future trajectories of direct human impacts on forests [i.e., deforestation/ degradation vs. conservation/ restoration, @ipcc_climate_2019;@ipcc_climate_2022], and also upon forest responses to climate change, which are likely to reduce the sink strength [@mcdowell_pervasive_2020].
<!--here I would definitely cite the Ruehr et al. 2023 paper-->

Accordingly, forests play a substantial role in international plans for climate change mitigation under the Paris Agreement [@unfccc_adoption_2015; @grassi_key_2017].<!--As of 2016, forest-based mitigation accounted for 26% of total planned greenhouse gas mitigation within Nationally Determined Contributions under the Paris Agreement [@grassi_key_2017].-->
Forest conservation, reforestation, and improved sustainable management all have significant -- and relatively cost-effective -- potential as climate change mitigation options [@roe_landbased_2021]. 
<!--SCP: You might get asked to say relative to what...the answer could be the social cost of carbon or more technological solutions. But it's probably fine to leave as is.-->
Conservation and reforestation have the fourth and fifth largest net emission reduction potentials or all mitigation options [@ipcc_summary_2022].
However, envisioned forest-based climate change mitigation initiatives do not always correspond to actual emission reductions implemented on the ground [e.g., @badgley_systematic_2022].
Realistic planning and reporting is critically needed to ensure that forest-based climate change mitigation initiatives are effective, and this requires solid scientific data and accounting frameworks [@anderson-teixeira_effective_2022; @deng_comparing_2021].

To this end, the International Panel on Climate Change (IPCC) provides guidance for national greenhouse gas inventories for reporting to the United Nations Framework Convention on Climate Change (UNFCCC) [@ipcc_2006_2006;@ipcc_2019_2019].
Under this guidance, greenhouse gas emissions to, or withdrawals from, the atmosphere are quantified on an annual basis for all managed land, which includes most of the world's forest land [@ogle_delineating_2018; @ellis_anthropogenic_2010].
The IPCC inventory guidelines include specific instructions for inventories for greenhouse gas (mainly CO~2~) exchanges between forest land and the atmosphere [`r appendix_IPCC_definitions`, @ipcc_agriculture_2006; @ipcc_2019_2019].
A tiered approach is employed, where the lowest tier (Tier 1) represents the simplest approach and relies on default parameter values -- for example, forest carbon (C) stocks values by ecozone<!-- [@fao_global_2012]--> and forest age class derived as the average of published estimates [@ipcc_2019_2019;@rozendaal_aboveground_2022].
Tier 1 values have improved over the years as more data and methods have become available [@requenasuarez_estimating_2019; @rozendaal_aboveground_2022], but there remains room for improvement.
For example, following the 2019 release of the latest IPCC guidelines, it was revealed that IPCC's Tier 1 default failed to capture eight-fold variation of C accumulation in regrowth forests within ecozones [@cook-patton_mapping_2020] and that C stocks in mature African tropical montane forests were two-thirds higher than the IPCC Tier 1 values for these forests [@cuni-sanchez_high_2021].
High variability of forest C cycling within ecozones [e.g., @cook-patton_mapping_2020; @cuni-sanchez_high_2021] means that it is crucial for practitioners to have access to locally-specific information, when available.
This rapid evolution of scientific information on C cycling in forests is valuable for informing climate change mitigation efforts but requires improved mechanisms for communicating the latest information from scientific researchers to the practitioners who need reliable estimates for greenhouse gas mitigation planning. 

To improve data accessibility for preparing greenhouse gas estimates, the IPCC created the Emission Factor Database (EFDB; [https://www.ipcc-nggip.iges.or.jp/EFDB/main.php](https://www.ipcc-nggip.iges.or.jp/EFDB/main.php)), which is intended as a recognized library of emission factors and other parameters that can be used for estimating greenhouse gas emissions and removals.
The EFDB can be used both for efforts to tally a nation's intended or accomplished greenhouse gas reductions, or as a basis of comparison for external parties to evaluate these inventories. 
The EFDB encourages researchers to submit estimates of emission factors or other related parameters [e.g., C stocks, net annual increments, and annual fluxes for various pools, @ipcc_2006_2006;@ipcc_2019_2019] from peer-reviewed journal articles or other accepted sources for inclusion in the database. 
Tens of thousands of relevant forest carbon estimates have been published -- and continue to be published at an accelerating rate -- but are not readily accessible to the practitioners assembling national greenhouse gas inventories.
To contribute to the goal of making forest C parameters available for accounting under IPCC guidelines, forest scientists need an accessible summary of EFDB's requirements and an efficient system for submission of data to the EFDB.

Our goal is to facilitate submission of forest C estimates from scientific studies to EFDB.
We document the process of submitting data to EFDB from the Global Forest Carbon Database, ForC ([https://forc-db.github.io/](https://forc-db.github.io/)), which is the largest collection of published estimates of forest C stocks, increments, and annual fluxes [@anderson-teixeira_forc_2018; @anderson-teixeira_carbon_2021; @anderson-teixeira_forcdb_2023], including data ingested from individual publications and relevant databases, including the Global Reforestation Opportunity Assessment (GROA) database [@cook-patton_mapping_2020, database doi: 10.5281/zenodo.3983644], and the Global Soil Respiration Database [SRDB-V5, @bond-lamberty_global_2010; @jian_restructured_2021].
We (1) map common scientific forest C estimation methods and definitions to those used by the IPCC;
(2) develop a semi-automated process for preparing ForC data for submission to EFDB;
and (3) assess the data in ForC relevant to EFDB and records that have been submitted to date.
We conclude with recommendations as to how the scientific community can better provide useful data for forest C inventories under the Paris Agreement.


# Materials and Methods {-}

Major steps for submission of data from ForC to EFDB included
(1) mapping ForC into EFDB, including aligning ForC terms and concepts with those defined by IPCC guidelines (summarized in `r appendix_IPCC_definitions`),
(2) revising ForC v3.0 to support semi-automated submissions to EFDB, yielding ForC v4.0 (detailed in `r appendix_ForC_updates`), and
(3) submitting data to EFDB. 

## 1. Mapping ForC to EFDB {-}

With input from the EFDB Technical Support Unit and referencing IPCC guidance [@ipcc_good_2003; @ipcc_2006_2006; @ipcc_2019_2019], we determined how EFDB fields should be populated using ForC fields (summarized in `r table_ForCfieldmapping`).
Fields in EFDB included several fields describing how the record fits within IPCC's framework (source/sink category, greenhouse gas type, C pool, relevant equations), several describing the C estimate itself (variable, value, units, 95% confidence limits), a few composite fields describing biotic and abiotic conditions (e.g., vegetation type, minimum diameter included, site location, climate, edaphic properties, notable disturbances, stand age, plot information), and a few describing the source (e.g., type, citation, data provider).
Most relevant ForC fields mapped directly into an EFDB field, either as the only contents of that field or as part of a composite record. <!--For example, ten ForC fields describing site location, climate, and edaphic properties all mapped into the EFDB field *Region/Regional conditions* (`r table_ForCfieldmapping`).-->
For some fields, simple conditional logic was used to populate EFDB fields based on ForC records.<!--For example, for stock variables presented in the original publication in units of dry organic matter mass (as opposed to C), several greenhouse gasses (CO~2~, CO, CH~4~, NO, NO~2~, N~2~O) were entered in the EFDB field indicating the greenhouse gases to which the record could be pertinent (*Gases* field) because these values could be used in calculations of greenhouse gas emissions from biomass burning [@ipcc_2006_2006]; otherwise, the only pertinent greenhouse gas would be CO~2~.--> 
For example, in cases where original studies did not present 95% confidence intervals (required by IPCC when available) but did present standard error or n and standard deviation, we calculated the 95% confidence intervals.<!--and populated the EFDB field with this information (noting the calculation in the EFDB field *Comments from Data Provider*)-->
There were two cases in which more complex mapping was required: (1) mapping of C cycle variables and (2) land classification.

### Carbon cycle variables {-}

Working in consultation with the EFDB Technical Support Unit and referencing IPCC guidance [@ipcc_good_2003; @ipcc_2006_2006; @ipcc_2019_2019], we identified ForC variables that were relevant to the IPCC methodology and EFDB (Fig. `r fig_variable_mapping`, `r appendix_IPCC_definitions`).
These included  organic matter or C stocks, net annual increments, influxes (a.k.a. "gross annual increments" by IPCC), and outfluxes for each IPCC-defied C pool (Table `r table_pools`, Fig. `r fig_variable_mapping`).
It is important to note that the correspondence of ForC variables to IPCC criteria often depends upon measurement protocols ("important sources of estimate variation" in Table `r table_pools`).
For example, ForC records of biomass and dead wood vary in the minimum stem diameter censused, such that some records would match the IPCC criteria whereas others would not.
Information on minimum diameters censused and other important sources of methodological variation were mapped into EFDB (`r table_ForCfieldmapping`).
Details on the mapping of ForC variables to EFDB are documented in the GitHub repository associated with this publication ([https://github.com/forc-db/IPCC-EFDB-integration](https://github.com/forc-db/IPCC-EFDB-integration)). 


\newpage
**Table `r table_pools`. IPCC-defined forest carbon pools with definitions and measurement methods.** <!--Definitions from IPCC Table 1.1. (See Table 1.1 in IPCC guidance).-->
```{r table_pools, eval = TRUE, echo=FALSE, warning=FALSE}
C_pools_table <- read.csv("figures_tables/C_pools.csv")
names(C_pools_table) <- stringr::str_replace_all(names(C_pools_table), "\\.", " ")

library(flextable)
set_flextable_defaults(font.family = "Arial", font.size = 8) # set font

ft <- flextable(C_pools_table)
ft <- autofit(ft)
ft <- width(ft, 1, width = 1)
ft <- width(ft, 2:4, width = 2)
ft <- bold(ft, i = 1, j = NULL, bold = TRUE, part = "header") # put title in bold
ft
```


![**Figure `r fig_variable_mapping`. Schematic illustrating the carbon pools defined under IPCC Guidelines for national greenhouse gas inventories; corresponding ForC variables, and relationships among them.** For each C pool, we show ForC variables corresponding to the stock, net annual increment, influx, and outflux. Most, but not all, of the `r n_relevant_variables` EFDB-relevant ForC variables are shown here. Correspondence of ForC variables to IPCC criteria often depends upon measurement protocols (e.g., minimum stem diameter censused). Additional caveats are as follows: (a,b) branch fall and mortality of stems below the minimum stem diameter censused, which are necessary for a full accounting of dead organic matter production but typically assumed negligible for calculations of biomass change, are excluded by common measurement practice (a) or ForC variable definition (b); (c) assumes that leaf production equals leaf fall, or that changes in foliage biomass are negligble; (d,e) belowground components excluded by common measurement practice (d) or ForC variable definition (e); (f) excludes movement of dead wood into litter through breakage or size reduction; (g) measurements often limited to litter layer (OL) and may exclude larger branches and stems classified as litter and/or the more decomposed layers of the O horizon. **This variable is techically EFDB-relevant but not selected for submission because their is no corresponding influx variable.](figures_tables/C_variable_mapping.png)




### Land classification {-}

Determination of the IPCC land-use category (i.e., Forest Land, Grassland, Wetlands, Cropland, Settlements, or Other Land; `r appendix_IPCC_definitions`) was made based on the dominant life form recorded in ForC.
Woody vegetation -- including early seral vegetation -- was classified as forest, consistent with the IPCC definition that Forest Land includes land expected to succeed to forest.
Mixtures of woody vegetation and grasses (i.e., anything from a shrub-encroached grassland to a tree-dominated savanna) were given dual classification of Forest Land and Grassland, indicating that records may be relevant to either category depending on the definition of forest applied (varies by country).
Grass- or crop-dominated ecosystems (included in ForC as controls for studies of forest regrowth following agricultural abandonment) were classified as Grassland and Cropland, respectively.

Classification into EFDB sub-categories was dependent upon stand age and site history.
For Forest Land ≥ 20 years old or of unknown (relatively mature) age, or Forest Land < 20 years old that was forest prior to a stand-clearing disturbance, the past land-use category was Forest Land, making the sub-category "Forest Land Remaining Forest land".
For forests <20 years old with history including cultivation/ tillage or grazing, past land-use categories were Cropland and Grassland, respectively, making land-use subcategories were "Cropland converted to Forest Land" and "Grassland converted to Forest Land", respectively.
For forests <20 years old with unspecified previous agricultural use, we assigned the sub-category "Land Converted to Forest land".
Forests <20 years old with unknown land use prior to the study date were simply classified as "Forest Land".
The same logic was applied for savannas, but including both Forest Land and Grassland as potentially relevant categories. 

Given the lack of public information needed to determine whether lands are classified as managed [@ogle_delineating_2018; @deng_comparing_2021], and because the IPCC's definition of managed land is more expansive than is commonly applied in the scientific literature and hence in ForC, we did not include any classification of land management status from ForC in the records submitted to EFDB.
However, we did provide auxiliary information that should be useful in making this determination, including geographical location and notable disturbance events. 

## 2. Updating ForC {-}

Previous versions of ForC [@anderson-teixeira_carbon_2016; @anderson-teixeira_forc_2018; @anderson-teixeira_carbon_2021] contained most of the information required by EFDB, and, more broadly, to inform C stock change calculations under the Paris Agreement. 
However, modest changes to the structure and contents of ForC were needed in order to provide all information required by EFDB and to improve ForC's capacity to serve as a repository of valuable information for forest C inventories under IPCC guidelines.
To support export of data to EFDB, and to improve the overall quality of the ForC database, we added or modified `r n_fields_modified` fields (`r table_ForCchanges`), defined `r n_new_variables` new variables, implemented enhanced quality control, manually reviewed >`r n_records_checked` records to obtain additional required information, and added `r n_records_added` new records.
Having implemented these changes, which are described in detail in `r appendix_ForC_updates`, to ForC v3.0 [@anderson-teixeira_carbon_2021], we released a new major version: ForC v4.0. 


## 3. Submission of ForC data to EFDB {-}

To submit complete, reviewed ForC records into EFDB, we created R scripts to restructure ForC records and populate EFDB's bulk import form (publicly available; see **Data Availability Statement**).
Criteria for data submission were that
(1) records had been checked against the original study and determined to be complete and correct, and as originally presented,
(2) all relevant records within the publication were included in ForC,
(3) the original study presented values in tables or text, as opposed to the values having been digitized from graphs or calculated based on related variables,
and (4) the records had not previously been submitted to EFDB.
Once converted into EFDB format, the records were reviewed and then sent to the IPCC's Technical Support Unit for submission to EFDB. 




# Results {-}

As of `r format(Sys.time(), '%B %d, %Y')`<!--inserts date upon knit-->, we had submitted only a fraction (~`r round(n_records_sent_to_EFDB/n_ind_records_EFDB_variables,2)*100`%) of relevant ForC records to EFDB, but these already comprised a substantial portion (`r percent_forest_records`%) of EFDB records for forest land.
Specifically, of the `r n_records_ForC` records in ForC (v4.0), ~`r n_ind_records_EFDB_variables` were independent records of EFDB-relevant variables (Fig. `r fig_variable_mapping`,  `r table_variables`).
We reviewed `r n_records_checked` records for submission to EFDB, of which
`r round((n_records_checked -(n_records_sent_to_EFDB - n_records_added))/n_records_checked,2)*100`% were inappropriate, mostly because values were digitized from graphs or calculated from related variables rather than directly presented.
We also added `r n_records_added` new records, mostly because EFDB required that all relevant records in a publication be submitted together. 
Of `r n_records_sent_to_EFDB` submitted records, EFDB's review panel accepted and posted `r n_records_posted_EFDB` but deemed `r round((1-(n_records_posted_EFDB/n_records_sent_to_EFDB)),2)*100`% not applicable.
Reasons included -- but were not limited to -- non-applicability to the IPCC methodology of the variable submitted (e.g., net ecosystem CO~2~ exchange, litter - OL layer; subsequently excluded from lists of relevant variables and counts of relevant records), inadequate information on the quality of data and thus on its replicability (e.g., confidence interval/uncertainty), and imprecisely described disturbance histories (e.g., "moderately"/ "severely" burned).
An iterative process of data submission review by the EFDB Technical Support Unit and review panel greatly improved but did not perfect (100% acceptance rate) our process.


ForC v4.0 records -- and the subset submitted to EFDB -- were unevenly distributed across all forested continents, biomes, and forest types (Figs. `r fig_map`-`r fig_histograms`).
Relative to forested area, submitted data under-represented the tropics and over-represented temperate regions, mirroring the composition of ForC (Figs. `r fig_map`, `r fig_histograms`b).
Submitted data covered all forested climates with substantial land area except the colder boreal regions (boreal tundra woodland; Figs. `r fig_map`, `r fig_climate`) and were concentrated in climates that are relatively common over Earth's land area (Fig. `r fig_climate`).
Among the records submitted to EFDB, the largest number came from North America, followed by Asia, South America, and Africa, with strongly disproportionate representation of North America and under-representation of South America, Africa, Europe, Australia, and Oceania (Fig. `r fig_histograms`c).
In terms of FAO ecozones, boreal coniferous forest, temperate continental forest, and temperate mountain systems had the most records and were the best represented relative to their global area, whereas other ecozones were poorly represented relative to their global area (Fig. `r fig_histograms`b).
By far the most records and best representation relative to global area were for needleleaf evergreen forests, followed by broadleaf deciduous and broadleaf evergreen (Fig. `r fig_histograms`b). 
The largest number of submitted records came from mature forests (>100 years), but young (<20 year) stands were best-represented relative to their global area, while 20-100 year stands were under-represented (Fig. `r fig_histograms`d).
ForC contains the records needed to provide more balanced geographical and forest type representation via selective data submission (Figs. `r fig_map`-`r fig_histograms`).

<!--In ForC v4.0, the largest number of records came from Asia, followed by North America, South America, and Europe, with relatively few records from Africa, Australia, and Oceania (Fig. `r fig_histograms`c).
Categorized by FAO ecozone, the greatest numbers of records came from subtropical humid forests, temperate mountain systems, and tropical rain forests, each with >2,000 independent records (Fig. `r fig_histograms`b). 
Boreal coniferous forests, temperate continental forests, subtropical mountain systems, and tropical moist deciduous forests had >1,000 independent records each, while other ecozones all had <1,000 records.
The most widely represented forest type was needleleaf evergreen, followed by broadleaf deciduous and broadleaf evergreen (Fig. `r fig_histograms`a).
In terms of stand age, the most represented age class was 20-100 years, followed by <20 years and then >100 years (Fig. `r fig_histograms`d).-->

![**Figure `r fig_map`. Map of sites in ForC shaded by number of independent records relevant (circles) and submitted (triangles) to EFDB.**  Underlying map shows FAO ecozones, and symbols are colored according to the number of records at each site. Inset stacked bar chart shows proportional representation by FAO ecozone for tree cover area (from SYNMAP, Jung *et al*, 2006), relevant ForC records, and submitted records. FAO ecozones (obtained from FAO’s GEOnetwork: http://www.fao.org:80/geonetwork) are coded as follows: Ba-Boreal coniferous forest, Bb-Boreal tundra woodland, BM-Boreal mountain systems, P-Polar, SBSh-Subtropical steppe, SBWh-Subtropical desert, SCf-Subtropical humid forest, SCs-Subtropical dry forest, SM-Subtropical mountain systems, TAr-Tropical rain forest,  TAwa-Tropical moist deciduous forest, TAwb-Tropical dry forest, TBSh-Tropical shrubland, TBWh-Tropical desert, TeBSk-Temperate steppe, TeBWk-Temperate desert, TeDc-Temperate continental forest, TeDo-Temperate oceanic forest, TeM-Temperate mountain systems, TM-Tropical mountain systems.](figures_tables/World_Map_of_sites_with_FAO_and_IPCC_data_sent.png)

![**Figure `r fig_climate`. Distribution of ForC records relevant (dots) and submitted (triangles) to EFDB within the global climate space of mean annual temperature (MAT) and mean annual precipitation (MAP).** Climate data are from the CRU TS3.10 dataset (0.5° resolution, 1990-2014, Harris *et al*, 2014). Background colors indicate the global land area with each MAT-MAP combination.](figures_tables/Climate_of_forC_sites_IPCC.png)

![**Figure `r fig_histograms`. Histograms of number of independent records in ForC relevant (grey) and submitted (black) to EFDB (both left axis), along with the relative global area of each type (red, right axis), organized by (a) dominant vegetation type, (b) FAO ecozone, (c) continent, and (d) stand age.** For dominant vegetation (a), 'Other' includes deciduous needleleaf, mixed broadleaf- needleleaf, non-woody vegetation (e.g., early successional), and incompletely classified or mixed forest types. Global coverage of each was obtained from SYNMAP (Jung *et al*, 2006). For FAO ecozones (b), codes are as listed in the caption of Figure `r fig_map`. The relative area of forests by age class was obtained from the global forest age database of Besnard *et al.* (2021, 2024).  ](figures_tables/Histogram_n_Relevant_and_Transferred_Records.png)

In terms of variables, ForC v4.0 contained records for `r n_relevant_variables_with_data` of the `r n_relevant_variables` variables (or closely-related variable groups) relevant to EFDB (Fig. `r fig_variable_mapping`,  `r table_variables`).
The records submitted to EFDB (`r n_relevant_variables_records_sent` variables) were very unevenly distributed across variables, mirroring the composition of ForC (`r table_variables`). 
The majority (`r round(n_records_sent_stocks/n_records_sent_to_EFDB, 2)*100`%) of records submitted were for C stocks, including
`r round(n_records_sent_biomass/n_records_sent_to_EFDB,2)*100`% for total biomass, 
`r round(n_records_sent_agb/n_records_sent_to_EFDB,2)*100`% for aboveground biomass,
`r round(n_records_sent_agb_components/n_records_sent_to_EFDB,2)*100`% for components of aboveground biomass (wood or foliage), 
`r round((n_records_sent_rootbiomass+n_records_sent_rootbiomass_components)/n_records_sent_to_EFDB,2)*100`% for root biomass or its components, 
`r round((n_records_sent_deadwood+n_records_sent_deadwood_components)/n_records_sent_to_EFDB,2)*100`% for dead wood or its components, 
`r round(n_records_sent_litter/n_records_sent_to_EFDB,2)*100`% for litter (O horizon),
and `r round(n_records_sent_soil/n_records_sent_to_EFDB,2)*100`% for soils.
Increment records (mostly for aboveground biomass) totaled `r round(n_records_sent_increments/n_records_sent_to_EFDB*100, 0)`% of records submitted.
The remaining `r round(n_records_sent_fluxes/n_records_sent_to_EFDB*100, 0)`% of records submitted described fluxes, all of which were either inputs or outputs to the aboveground biomass pool, with some also describing inputs to dead wood or litter pools (Fig. `r fig_variable_mapping`,`r table_variables`).
Many of the EFDB-relevant variables remain poorly represented in ForC (`r table_variables`).



# Discussion {-}

Here we have developed a framework for submitting records from ForC v4.0, the largest compilation of ground-based forest C estimates, to the EFDB, thus making those data more accessible for reporting CO~2~ emissions and removals from forest land consistent with good practice in the IPCC guidelines  [@ipcc_2006_2006;@ipcc_2019_2019].
As of `r format(Sys.time(), '%B %d, %Y')`, `r n_records_posted_EFDB` ForC records have been posted to EFDB, which represents just `r round(n_records_posted_EFDB/n_ind_records_EFDB_variables*100,0)`% of potentially relevant records in ForC v4.0, but currently comprises `r percent_forest_records`% of the total EFDB records for forest land.
The records present in ForC and submitted to EFDB are very unevenly distributed across variables, regions, and forest types (Figs. `r fig_map`-`r fig_histograms`, `r table_variables`), reflecting broader patterns in allocation of research effort and pointing to scientific research and reporting needs.

Based on our experience contributing forest C data to EFDB via ForC, we make several recommendations as to how scientists can improve forest C records in EFDB through database work, new data collection and analysis, and reporting.

### Database needs {-}

There is vast potential to expand forest C data in EFDB by completing the process of reviewing and submitting data that are already in ForC (Figs. `r fig_map`-`r fig_histograms`).
So far, only ~`r round(n_records_sent_to_EFDB/n_ind_records_EFDB_variables,2)*100`% of the EFDB-relevant data in ForC have been submitted to EFDB.
Although this process requires manual review of records, the submission of new records to EFDB is greatly facilitated by the fact that most pertinent information for each record is already entered in ForC and can be easily prepared for submission to EFDB using the system developed here.
Future efforts to review studies for submission should optimize for representation across geographic regions, forest types, and variables, giving priority to those from currently under-represented regions and forest types (Figs. `r fig_map`-`r fig_histograms`, `r table_variables`). Other categories of records to prioritize include those from countries relying on existing data for their greenhouse gas inventories (Tier 1 or 2 methodology), the variables most needed by EFDB users, and the most up-to-date records.

In addition to the large potential to expand EFDB using records already in ForC, there are extensive EFDB-relevant forest C data that are not currently included in ForC, with more being published on a nearly daily basis. 
Coverage of particular variables or regions could be vastly improved through systematic review of the literature, although this requires focused and extensive manual effort.
Recent efforts have compiled large databases of relevant data from monoculture plantation forests [@bukoski_rates_2022] and mixed species plantation forests [@feng_multispecies_2022;@warner_higher_2022], and such a compilation is in works for agroforestry (Susan Cook-Patton, unpublished data). 
Beyond expanding collections of relevant forest C records, such reviews are valuable for assessing the availability of published records and identifying variables and regions that require additional data collection and analysis. 

### Data collection and analysis needs {-}

New data collection and analyses are needed to fill notable knowledge gaps.
While aboveground biomass stocks in particular have received -- and continue to receive -- by far the most research attention [`r table_variables`,@nisar_nasaisro_2018; @anderson-teixeira_carbon_2021;@dubayah_global_2020; @quegan_european_2019], production of an accurate global map of forest C stocks remains an ongoing challenge [@araza_decade_2023]. 
Other pools and variables remain poorly quantified and highly uncertain for many parts of the world [`r table_variables`,@tifafi_large_2018; @anderson-teixeira_carbon_2021], introducing substantive uncertainties into global forest C budgets [@pan_large_2011;@harris_global_2021].
Furthermore, data distribution is very uneven across forest types and geographical regions (Figs. `r fig_map`-`r fig_histograms`).
For instance, data on C cycling of tropical forests -- particularly in Africa -- remains relatively sparse, in large part due to substantial barriers to data collection and distribution [@delima_making_2022].
More generally, belowground C measurements remain sparse globally.
Significant investment in research and researchers focused on ground-based measurement of forest C in such regions will be important to filling knowledge gaps in forest C cycling [@delima_making_2022;@araza_decade_2023; @labrière_forest_2023].

Several EFDB-relevant variables have not been calculated and presented as frequently as would be possible given existing forest census data and minimal extra research effort.
For example, aboveground woody mortality and aboveground biomass increment can be calculated from the same census data as aboveground woody productivity, yet the latter has received far more research attention [`r table_variables`, @anderson-teixeira_carbon_2021]. 
Similarly, live coarse root biomass, total biomass, and changes therein could in theory be estimated in parallel with aboveground biomass, with the greatest barrier being that allometric models for estimating root biomass are not as reliable or easily available as are those for aboveground biomass [@chave_improved_2014;@rejou-mechain_biomass_2017;@gonzalez-akre_allodb_2022].
However, while equations for estimating root (and thereby total) biomass require improvement, they do exist for many forest types [e.g., @brassard_coarse_2011; @chojnacky_updated_2014;  @waring_overlooking_2017; @mokany_critical_2006].
We recommend that, when possible, researchers calculate and report all EFDB-relevant variables.


### Data reporting needs {-}

We recommend that, in order to make research most useful for estimating C stock changes following IPCC guidelines, researchers calculate and report results according to IPCC good practice (Table `r table_recommendations`).
Importantly, simple decisions on the presentation of results will determine whether the records meet the current criteria for inclusion in EFDB.
For example:
(1) presenting data only in a figure makes them ineligible for inclusion in EFDB, whereas presentation in a table or supplementary data file allows inclusion while supporting FAIR goals [@stall_make_2019];
(2) direct presentation of all relevant variables allows inclusion, whereas presenting only components of variables of interest (e.g., parsing litter into fine woody debris, OL, OF, and OH layers) or requiring simple mathematical operations to obtain a variable of interest (e.g., *delta.agb* = *ANPP_woody* - *woody.mortality.agb*) disqualifies records from inclusion;
(3) matching IPCC-defined thresholds for defining C pools (Table `r table_pools`), which may vary by country, can make the data far more relevant estimating forest C stock changes according to IPCC guidelines (e.g., using a 10 cm cutoff between dead wood and litter, presenting soil C to a depth of 30 cm).
It should also be emphasized that reporting of 95% confidence intervals (or other metrics of error), when applicable, is highly desirable and makes the data more relevant to IPCC. 

\newpage
**Table `r table_recommendations`. Recommended best practices for reporting forest C estimates of value to national greenhouse gas inventories under IPCC guidance.**
```{r table_recommendations, eval = TRUE, echo=FALSE, warning=FALSE}
recommendations_table <- read.csv("figures_tables/recommendations.csv")
names(recommendations_table) <- stringr::str_replace_all(names(recommendations_table), "\\.", " ")

library(flextable)
set_flextable_defaults(font.family = "Arial", font.size = 8) # set font

ft <- flextable(recommendations_table)
ft <- autofit(ft)
ft <- width(ft, 1, width = 1)
ft <- width(ft, 2:3, width = 3)
ft <- bold(ft, i = 1, j = NULL, bold = TRUE, part = "header") # put title in bold
ft
```

For those compiling published records (e.g., for meta-analyses), the data set can have added value if all information required by EFDB is extracted from original publications.
<!--SCP - optional addition : Realistic  planning and reporting is critically needed to ensure that forest-based climate change mitigation initiatives are effective, and this requires solid scientific data -->
This includes -- but is not limited to -- retaining original values as presented without modification or rounding, noting whether data were digitized, recording confidence intervals, and recording all required fields (as indicated in the EFDB's bulk import template).
The significant effort required to map a database into EFDB has been accomplished here (`r table_ForCfieldmapping`), and we hope that it will prove useful as a model for other efforts.


# Conclusions {-}

As human society strives to achieve net-zero greenhouse gas emissions, forest researchers can make their research more useful for forest C inventories under IPCC guidelines by calculating and reporting results in ways that are consistent with methodologies provided in the IPCC guidelines (Tables `r table_pools`, `r table_recommendations`).
In addition, substantial investments in research and researchers focused on ground-based measurement of forest C will be required to fill knowledge gaps and thereby increase the accuracy of forest CO~2~ inventories for forest lands under the Paris Agreement. 
This challenge is heightened by the fact that forests are changing rapidly [e.g., @mcdowell_pervasive_2020], and data collected a decade or more in the past may be increasingly less accurate.
This heightens the need for an efficient system of making forest C data accessible for national greenhouse gas inventories.
We view the system developed here for submitting ForC data to the IPCC EFDB as one important step towards that goal.


# Acknowledgements {-}
We gratefully acknowledge the substantial contributions of Valentyna Slivinska and Sandro Federici for collaboration on the conception, design, and technical review of this project.
Thank you to all researchers who collected and published the data contained in ForC, and to all research assistants and collaborators who have contributed to building the database.
Thank you to Avni Malhotra for helpful comments on an earlier draft of this manuscript.
Funding for this study was provided by the Smithsonian (Forest Global Earth Observatory, Smithsonian Working Land and Seascapes) and a Bezos Earth Fund grant to the Nature Conservancy with a sub-grant to NZCBI.

# Author Contribution {-}
KAT and VH conceived and designed the project; VH wrote the scripts for database management, data submission to EFDB, and the analyses presented here; MW, TR, and RBM added and reviewed ForC data, BBL and SCP contributed large databases to ForC (EFDB and GROA, respectively); CP provided methodological expertise; KAT, VH, and MW prepared the first draft of the manuscript; all authors reviewed the results and approved the final version of the manuscript.
 
# Data Availability Statement  {-}
All code and data are openly available. The ForC database and associated code are available via the ForC repository within the ForC-db organization on GitHub (https://github.com/forc-db/ForC), and the version used here (ForC v4.0) is archived in Zenodo [@anderson-teixeira_forcdb_2023, DOI: 10.5281/zenodo.8020861]. The data and code associated with data submission to EFDB and preparation of this manuscript are available via the the IPCC-EFDB-integration repository within the ForC-db organization on GitHub  (https://github.com/forc-db/IPCC-EFDB-integration) and archived in Zenodo (DOI: 10.5281/zenodo.8021474). 

# Supplementary Information {-}

`r table_ForCchanges`.  Updates to ForC field implemented between releases of v3.0 and v4.0

`r table_ForCfieldmapping`. Mapping of ForC fields to EFDB

`r table_variables`. Numbers of ForC records and EFDB submissions by variable

`r appendix_IPCC_definitions`. Primer on forest land classification and carbon pools under IPCC guidelines

`r appendix_ForC_updates`. Updates to ForC (ForC v4.0)



\clearpage
# References {-}

