---
title: ''
output: 
    pdf_document
bibliography: references.bib
csl: new-phytologist.csl

---


```{r eval = TRUE, echo=FALSE, warning=FALSE, results='hide'}


#Appendices
appendix_ForC_updates = "Methods S1"
appendix_IPCC_definitions = "Notes S1"
table_ForCchanges = "Table S1"
table_ForCfieldmapping = "Table S2"
table_variables ="Table S3"


```



**`r table_ForCchanges`. Updates to ForC field implemented between releases of v3.0 and v4.0.** 
```{r table_ForCchanges, eval = TRUE, echo=FALSE, warning=FALSE, message = FALSE}
library (readr)
urlfile="https://raw.githubusercontent.com/forc-db/ForC/master/database_management_records/record_of_changes.csv"
record_of_changes <-read_csv(url(urlfile))
names(record_of_changes) <- stringr::str_replace_all(names(record_of_changes), "\\.", " ")
record_of_changes <- record_of_changes[c(1:5)] #include all fields but date of change
#record_of_changes<- record_of_changes[which(!grep("housekeeping",record_of_changes$Motivation)) , c(1:5)]  #trying to exclude rows for which Motivation = "housekeeping", but this is not working
record_of_changes[is.na(record_of_changes)]<- "" #replace NAs with blanks

library(knitr)
library(kableExtra)
knitr::kable(record_of_changes, format = "latex",
             longtable = TRUE) %>%
  kableExtra::kable_styling(full_width = TRUE, latex_options = c("repeat_header")) %>%
  kable_styling(font_size = 8) %>%
  row_spec(0, bold = TRUE)
```

\clearpage


**`r table_ForCfieldmapping`. Mapping of ForC fields to EFDB.** Details documented in the public GitHub repository associated with the project, IPCC-EFDB-integration repository within the ForC-db organization (file *ForC-EFDB_mapping.csv* available at [https://github.com/forc-db/IPCC-EFDB-integration/blob/main/doc/ForC-EFDB_mapping/ForC-EFDB_mapping.csv](https://github.com/forc-db/IPCC-EFDB-integration/blob/main/doc/ForC-EFDB_mapping/ForC-EFDB_mapping.csv)). 
```{r table_ForCfieldmapping, eval = TRUE, echo=FALSE, warning=FALSE}
library(knitr)
library(kableExtra)

ForC_variables_included <- read.csv("figures_tables/ForC_fields_included.csv")
names(ForC_variables_included) <- stringr::str_replace_all(names(ForC_variables_included), "\\.", " ")
column_names <- names(ForC_variables_included) # Define column names
column_names[5] <- "Required*" # Add asterisk for footnote
names(ForC_variables_included) <- column_names# Set column names of the matrix
ForC_variables_included <- ForC_variables_included[c(1:5)] 

knitr::kable(ForC_variables_included, format = "latex",
             longtable = TRUE) %>%
  row_spec(0, bold = TRUE) %>%
  kable_styling(font_size = 8) %>%
kableExtra::kable_styling(full_width = TRUE, latex_options = c("repeat_header", "hold position"))

```
\* *Required* field indicates whether the field is required by EFDB: yes = value required; (yes) = input required, missing value acceptable if not reported; auto = present within ForC infrasructure, and therefore will always be exported to EFDB ; (no) = not required for EFDB, but required for ForC and therefore will always be exported to EFDB; no = not required, but exported to EFDB when a value is present.

\clearpage


\newpage
**`r table_variables`. Numbers of records of ForC variables (or closely related variable groups) relevant to, and sent to, EFDB.**
```{r table_variables, eval = TRUE, echo=FALSE, warning=FALSE}
C_variables_table <- read.csv("figures_tables/C_variables.csv")
names(C_variables_table) <- stringr::str_replace_all(names(C_variables_table), "\\.", " ")
C_variables_table[is.na(C_variables_table)]<- "" #replace NAs with blanks

library(flextable)
set_flextable_defaults(font.family = "Arial", font.size = 8) # set font

ft <- flextable(C_variables_table[,-1])
ft <- autofit(ft)
ft <- width(ft, 1, width = 2)
ft <- width(ft, 2:5, width = 1)
ft <- bold(ft, i = 1, j = NULL, bold = TRUE, part = "header") # put title in bold
ft <- bold(ft, i = which(C_variables_table$bold), j = NULL, bold = TRUE, part = "body") # put categories in bold
ft
```
